### 2016-12-08 tex/pres.tex
my portion of the presentation is complete

\documentclass[10pt, xcolor={dvipsnames}, aspectratio=169]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[super, square]{natbib}
\usepackage{mathtools, relsize, graphicx, amssymb, amsthm}
\usepackage{listings}
\usepackage{color}
\usepackage[strings]{underscore}

\lstset{
    frame=single,
    basicstyle=\footnotesize,
    keywordstyle=\color{purple},
    numbers=left,
    numbersep=5pt,
    showstringspaces=false, 
    stringstyle=\color{blue},
    tabsize=4,
    language=C++
}

\mode<presentation>{\usetheme{Berlin} \usecolortheme{beetle}}

\title{Parallel Sorting of Roughly-Sorted Sequences}
\author{Anthony Pfaff, Jason Treadwell}
\institute{CSCI 5172 $|$ CU Denver $|$ Fall '16}
\date{12.10.2016}

\begin{document}
\setbeamertemplate{navigation symbols}{} %remove navigation symbols
\bibliographystyle{plain}
\nocite{*}
% gets rid of the references button
\renewcommand{\bibsection}{\subsubsection*{\bibname}}
\graphicspath{{./}}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\transfade
\frametitle{Introduction}
Sorting a sequence according to some ordering is a foundational problem of computer science. \newline

Sorting has crucial and non-obvious applications. There exist a great many such algorithms. \newline

The optimal runtime of sorting a sequence of length $n$ by \textit{comparison} is $\Theta(n \lg n)$.
\end{frame}

\begin{frame}
\frametitle{Sort Algorithms}
The comparison sorts include Mergesort, Heapsort, Quicksort, etc.\ and have diverse strengths. \newline

The former two are asymptotically optimal; Quicksort frequently outruns them, but is $O(n^2)$. \newline

We can beat the linearithmic runtime bounds through additional analysis of the input array.
\end{frame}

\begin{frame}
\frametitle{Roughly-Sorted Sequences}
The sequence $A = \{a_0, a_1, \cdots, a_{n-1}\}$ is $k$-sorted if it satisfies
$$a_i \leq a_j \quad\forall\; i < j - k\,,\quad 0 \leq i \leq j < n \,.$$ \newline

The radius of $A$ is the smallest $k$ such that $A$ is $k$-sorted:
$$\text{radius}(A) = \max\{j - i \mid j > i,\, a_i > a_j \} \,.$$
\end{frame}

\begin{frame}
\frametitle{Conventions}
We'll only consider sequences of 32-bit integers backed by \textit{random-access arrays}. \newline

The array $A = \{a_0, a_1, \cdots, a_{n-1}\}$ is \textit{sorted} if its elements are all ordered in nondecreasing order. \newline

When we say $A$ is $k$-sorted, we imply that $k$ is minimal (i.e., the radius of $A$ is $k$).
\end{frame}

\begin{frame}
\frametitle{Roughly-Sorted Sequences (cont.)}
Some applications involve sorted arrays that become slightly perturbed. \newline

We might be able to exploit their partial sortedness to beat the $\Omega(n \lg n)$ lower runtime bound. \newline

Comparison-based sorts are prone to divide-and-conquer and, therefore, opportunities for parallel speedup.
\end{frame}

\begin{frame}
\frametitle{Roughsort}
Roughsort exploits the presortedness of the array $A$ by applying a radius \textit{halving} algorithm. \newline

The algorithm sorts $A$ by halving its radius $\lg k$ times in runtime $\Theta(n \lg k) = O(n \lg n)$. \newline

Roughsort does invite parallel speedup, yet must determine $k$ to effectively sort $A$.
\end{frame}

\begin{frame}
\frametitle{Determining the Radius}
Using min/max prefix scans, we compute the \textit{characteristic sequences} of $A$:
$$LR(A) = \{b_i\}\,,\quad RL(A) = \{c_i\}\,,\quad 0 \leq i < n\,,$$
$$b_i = \max\{a_0, a_1, \cdots, a_i\}\,,\quad c_i = \min\{a_i, a_{i+1}, \cdots, a_{n-1}\}\,.$$

The radius $k$ of $A$ is the maximum element from the \textit{disorder measure} of $A$:
$$DM(A) = \{d_i\}\,,\quad d_i = \max\big\{\{i - j \mid c_i < b_j\} \cup \{0\} \big\}\,,\quad k = \max DM(A) \,.$$

Finding $k$ thus takes linear time and space, preserving the $\Theta(n \lg k)$ complexity of Roughsort.
\end{frame}

\begin{frame}
\frametitle{Halving the Radius}
\begin{enumerate}
\item Partition each consecutive run of $k$ elements in $A$ about the mean of the run.
\item Starting at element $a_{\lfloor k/2 \rfloor}$ to stagger the partitions, repeat step 1 and go to 3.
\item Repeat step 1 and halt.
\end{enumerate}

Each partition is performed using STL's \texttt{nth_element()} and takes linear time. \newline

Our sequential implementation thus sorts $A$ in place by halving its radius $\lg k$ times, taking linear space and
  $\Theta(3(n/k \cdot k) \lg k) = \Theta(n \lg k)$ time.
\end{frame}

% BEGIN CUDA INTRO
\begin{frame}
\frametitle{}
\end{frame}

\begin{frame}
\frametitle{}
\end{frame}
% END CUDA INTRO

% BEGIN PAR RADIUS
\begin{frame}
\frametitle{}
\end{frame}

\begin{frame}
\frametitle{}
\end{frame}
% END PAR RADIUS

% BEGIN PAR RADIUS RESULTS
\begin{frame}
\frametitle{}
\end{frame}

\begin{frame}
\frametitle{}
\end{frame}
% END PAR RADIUS RESULTS

\begin{frame}
\frametitle{Parallel Roughsort: Failed Approach}
Our first implementation simply parallelized the halving algorithm of our sequential Roughsort. \newline

We replaced \texttt{nth_element()} with Thrust's \texttt{sort()}, an optimized radix sort. \newline

We tried to launch every partitioning in parallel using \textit{streams} and got abysmal results.
\end{frame}

\begin{frame}
\frametitle{Parallel Roughsort: Better Approach}
\begin{enumerate}\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}
\item In parallel, \textbf{sort} all consecutive runs of $2k$ elements in $A$.
\item Repeat step 1, but begin the first run at $a_k$; halt.
\end{enumerate}
This process sorts each segment and needs only a single iteration to fully sort $A$. \newline

We implemented it as a CUDA kernel where the array segments were divided up among many parallel threads,
  but each thread sequentially sorted its designated array segment.
\end{frame}

\begin{frame}
\frametitle{Testing the Implementation}
We tested our Roughsort implementations against sequential and parallel Mergesort as well as against Bubblesort. \newline

Each test generated a random $k$-sorted array for some given $k$ and length $n$, then fed it to all five algorithms, averaging
  their results over eight runs. Each array was sourced by a hardware RNG and sorted before being
  $k$-perturbed and shuffled. \newline

The tests were run on a workstation with a high-end, quad-core Xeon CPU as well as a high-end, consumer-grade
  Nvidia GeForce 1080 GTX GPU suitable for workloads over 32-bit array elements.
\end{frame}

\begin{frame}
\frametitle{Sort Runtimes $(k = 2)$}
\begin{figure}
\scalebox{0.4}{\input{plots/k2}}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Sort Runtimes $(k = 15)$}
\begin{figure}
\scalebox{0.4}{\input{plots/k15}}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Sort Runtimes $(n = 0.75 \cdot 10^6)$}
\begin{figure}
\scalebox{0.4}{\input{plots/n750k}}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Sort Runtimes $(n = 1.25 \cdot 10^6)$}
\begin{figure}
\scalebox{0.4}{\input{plots/n1250k}}
\end{figure}
\end{frame}

% BEGIN EXPLANATION OF BAD PAR. PERFORMANCE
\begin{frame}
\frametitle{}
\end{frame}

\begin{frame}
\frametitle{}
\end{frame}
% END EXPLANATION OF BAD PAR. PERFORMANCE

\begin{frame}
\frametitle{Conclusion and Further Research}
\end{frame}

\end{document}